### 2019-07-05 kindly/sdtyping/extra.tex
Extend/reorganize the syntax-directed typing section.

\section{Annotating regions}
\label{regionannot}

So far, all \lang programs have been fully annotated with regions information.
We now show how to infer these regions annotations based on
optionally-annotated programs.
First, we extend the region annotation to $\region{S}{E}$ where $S$ is
a set of variables. This annotation, defined below, is equivalent to nested
region annotations for each individual variable.

\begin{align*}
  \region{x;S}{e} &= \region{x}{\region{S}{e}}& \region{\emptyset}{e} &= e\\
\end{align*}

\Cref{fig:region-annotation} define a rewriting relation $\RannotT{e}{e'}$
which indicates that an optionally annotated term $e$ can be written
in a fully annotated term $e'$.
Through the rule \textsc{Rewrite-Top}, this is defined
in term of an inductively defined relation
$\Rannot{e}{e'}{S}$ where $n$ is the current nesting and $S$ is a set of
variable that are not enclosed in a region yet.
The base cases are constants, variables and borrows.
The general idea is to start from the leafs of the syntax tree, create a
region for each borrow, and enlarge the region as much as possible.
This is implemented by a depth-first walk of the syntax
tree which collects each variable that has a corresponding borrow.
At each step, it rewrites the inner subterms,
consider which borrow must be enclosed by a region now, and
return the others for later enclosing. Binders force immediate
enclosing of the bound variables, as demonstrated in rule \textsc{Rewrite-Lam}.
For nodes with multiple children, we
use a scope merge operator to decide if regions should be placed and where.
This is shown in rule \textsc{Rewrite-Pair}.
The merge operator, written $\getBorrows{B_l}{B_r}{(S_l,S,S_r)}$, takes
the sets $B_l$ and $B_r$ returned by rewriting the subterms
and returns three sets: $S_l$ and $S_r$ indicates the variables
that should be immediately enclosed by a region on the left and right
subterms and $S$ indicates the set of the yet-to-be-enclosed variables.
As an example, the rule \textsc{AnnotRegion-MutLeft} is applied
when there is an immutable borrow and a mutable borrow. In that case, a
region is created to enclose the immutable borrow, while the mutable
borrow is left to be closed later. This is coherent with the rules
for environment splitting and suspended bindings from \cref{sdtyping}.
%
Explicitly annotated regions are handled specially through
rule \textsc{Rewrite-Region}. In that case, we assume that all inner
borrows should be enclosed immediately.

\begin{figure*}[!hbt]
  \centering
  \input{region-annot}
  \caption{Automatic region annotation --- $\RannotT{e}{e'}$}
  \label{fig:region-annotation}
\end{figure*}

\section{Syntax directed typing}
\label{appendix:sdtyping}

\subsection{Constraint language}

\newcommand\A{\mathcal A}
\newcommand\SC{\mathcal S}

Let us note $\A$ our constraint system. The full grammar of constraints is
given in \cref{grammar:constraint}.
$\A$ is defined as the smallest cylindric term constraint system that
satisfies the axiom shown in \cref{rules:entail}.
We follows the traditional HM(X) formulation
with conjunctions, projections and type inequalities.
The new element specific to our approach are kind inequalities.
Entailment is noted $\entail{C}{D}$, where $D$ is a consequence of the
constraints $C$.
We say that $C$ and $D$ are equivalent, noted $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
\TODO{Give the cylindric properties ?}

\begin{figure}[tp]
  \centering
  \begin{align*}
    C &::= \Cleq{\tau_1}{\tau_2}
        \mid \Cleq{k_1}{k_2}
        \mid C_1 \Cand C_2
        \mid \Cproj{\tvar}{C}
        \mid \Cproj{\kvar}{C}
  \end{align*}
  \caption{The constraint language}
  \label{grammar:constraint}
\end{figure}

\begin{figure*}[tp]
  \input{infer/lattice}
  \caption{Lattice inequalities -- $k \lk_\Lat k'$}
  \input{infer/entails}
  \caption{Base entailment rules -- $\entail{C}{D}$ }
  \label{rules:entail}
\end{figure*}

We note $\SC$ the set of solved forms
which can be used inside type and kind schemes.
We define $\SC$ as $\A$ quotiented by the relation $\equivC$.
%
We consider the existence of a function $\operatorname{normalize}$ which takes
a constraint in $\A$ and a substitution $\psi$ and returns a constraint
in solved form $C' \in \SC$,
and an updated substitution. We detail the implementation
of the normalization function in \cref{sec:normalize}

% $\mathcal S$ is composed only of kind
% inequalities \emph{over variables}. For convenience, if $C\in\mathcal S$, we
% note $C$ as a list of kind inequalities: $\Cleq{\kvar_i}{\kvar_{i'}}^n$.
% \TODO{Extend the properties of solved forms}

\subsection{Kinding}

We note $\inferSK{C}{\E}{\tau}{k}$
when $\tau$ has kind $k$ in environment $\E$ under constraints $C$.
The rules are shown in \cref{rules:sd-kinding}.
Kinds and types follow a small calculus with variables ($\tvar$,\dots),
functions (type constructors $\T{t}$), application ($\tapp{t}{\Multi{\tau}}$)
and primitives such as types for arrows ($\tau\tarr{k}\tau'$) and
borrows ($\borrowty{k}{\tau}$).
Kind checking can thus be done in a fairly straightforward, syntax-directed
fashion by simply following
the syntax of the types. Kind arrows can only appear when looking
up the kind scheme of a type constructor $\T t$. Kind arrows are forbidden
in any other contexts.

\begin{figure}[ht]
  \centering
  \input{sdtyping/sd-kinds}
  \caption{Syntax-directed kinding rule --
    $\inferSK{C}{\E}{\tau}{k}$}
  \label{rules:sd-kinding}
\end{figure}

\subsection{Environments}

\begin{figure}[tp]
  \input{sdtyping/sd-splitting}
  \caption{Splitting --- environments $\lsplit
    C\E\E\E$; binders $\bsplit Cbbb$}
  \label{fig:sd-splitting}
\end{figure}

\begin{figure}[tp]
  \input{sdtyping/sd-borrowing}
  \caption{Borrowing --- environments $\lsplit
    C\E\E\E$; binders $\bsplit Cbbb$}
  \label{fig:sd-borrowing}
\end{figure}

\subsection{Typing}

\begin{figure*}[tp]
  \input{sdtyping/sd-environments}
  \caption{Rewriting constraints on environments and types}
  \label{fig:contraints-environments-types}
\end{figure*}
\begin{figure*}[tp]
  \input{sdtyping/sd-typing}
  \caption{Syntax-directed typing rules}
  \label{fig:syntax-directed-typing}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: