### 2019-07-07 kindly/infer/extra.tex
Select the various part of the type inference section.

\section{Inference rules}
\label{appendix:infer}

\begin{figure*}[htp]
  \input{infer/typing}
  \caption{Type inference algorithm -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
  \label{rules:typing}
\end{figure*}

\begin{figure}[tp]
  \centering
  \begin{mathpar}
    { \Weaken_{\bvar{x}{\sigma}}(\Sv) =
      \begin{dcases}
        \Ctrue& \text{if } x\in\Sv\\
        \Cleq{\sigma}{\kaff_\infty}&\\
      \end{dcases}
    }

    \inferrule{
      \inferK{C \Cand C_x} \E \tau {k'} \\
      D = C \Cand C_x \Cand \Cleq{k'}k
    }{
      \E \vdash \Cleq{\forall \kvar_i \forall (\tvar_j:k_j).\ \qual{C_x}{\tau}}{k} \Crewrite  D
    }
  \end{mathpar}
  \caption{Auxiliary rules for inference}
  \label{op:usgmap}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: