### 2019-07-10 kindly/infer/extra.tex
Add the proof for completeness.

\section{Type Inference}
\label{appendix:infer}

\begin{figure*}[p]
  \input{infer/typing}
  \caption{Type inference rules -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
  \label{rules:typing:full}
\end{figure*}

In this appendix, we provide the complete type inference rules
and show that our type inference algorithm is sound and complete.
The kinding and constraints rules are already shown in \cref{inference}.
The type inference rules are shown in \cref{rules:typing:full}.
The various theorems and their proofs are direct adaptations
of the equivalent statements for HM(X) \citep{sulzmann1997proofs}.

\clearpage
% \begin{figure}[tp]
%   \centering
%   \begin{mathpar}

%     % \inferrule{
%     %   \inferK{C \Cand C_x} \E \tau {k'} \\
%     %   D = C \Cand C_x \Cand \Cleq{k'}k
%     % }{
%     %   \E \vdash \Cleq{\forall \kvar_i \forall (\tvar_j:k_j).\ \qual{C_x}{\tau}}{k} \Crewrite  D
%     % }
%   \end{mathpar}
%   \caption{Auxiliary rules for inference}
%   \label{op:usgmap}
% \end{figure}

\input{infer/metatheory}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: