### 2019-07-11 kindly/nonanon.tex
Add the non anonymous materials.

\documentclass{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}

\title{Non-anonymous material for the article\\
  ``Kindly Bent to Free Us''}
\date{}

\begin{document}

\maketitle

An online playground for the type-checker is available at:
\begin{center}
  \url{https://drup.github.io/pl-experiments/affe/}
\end{center}

The implementation is available at
\begin{center}
  \url{https://github.com/Drup/pl-experiments}
\end{center}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: