### 2019-11-07 kindly/infer/extra.tex
Start improving the inference section.

\section{Type Inference}
\label{appendix:infer}

In this appendix, we provide the complete type inference rules
and show that our type inference algorithm is sound and complete.
The constraints rules are already shown in \cref{inference}.
Kind inference is presented in \cref{rules:kinding:full}.
The type inference rules are shown in \cref{rules:typing:full}.
The various theorems and their proofs are direct adaptations
of the equivalent statements for HM(X) \citep{sulzmann1997proofs}.

\subsection{Kind inference}
\label{rules:kinding:full}
%
\begin{figure}[tb]
  \centering
  \input{infer/kinds}
  \caption{Kind inference rules -- $\inferK{(C,\unif)}{{\bm\E}}{{\bm\tau}}{k}$}
  \label{rules:kinding}
\end{figure}
%
We note $\inferK{(C,\unif)}{{\bm\E}}{{\bm\tau}}{k}$ when type $\tau$ has kind $k$
in environment $\E$ under constraints $C$ and unifier $\unif$.
$\bm\E$ and $\bm\tau$ are the input parameters of
our inference procedure.
We present the kind inference algorithm as a set of rules in
\cref{rules:kinding}.
Higher-kinds are not generally supported
and can only appear by looking-up the kind scheme of a type constructor,
for use in the type application rule {\sc KApp}.
Type variables must be of a simple kind in rule {\sc KVar}.
Kind schemes are instantiated in the {\sc KVar} rules by creating
fresh kind variables and the associated substitution.
{\sc KArr} and {\sc KBorrow} simply returns the kind of the primitive
arrow and borrow types.
The $\operatorname{normalize}$ function is used every time several constraints
must be composed in order to simplify the constraint and return a most general
unifier.

\begin{figure*}[!btp]
  \input{infer/typing}
  \caption{Type inference rules --
    $\inferW{\Sigma}{(C,\psi)}{{\bm\E}}{{\bm e}}{\tau}$ }
  \label{rules:typing:full}
\end{figure*}

% \begin{figure}[tp]
%   \centering
%   \begin{mathpar}

%     % \inferrule{
%     %   \inferK{C \Cand C_x} \E \tau {k'} \\
%     %   D = C \Cand C_x \Cand \Cleq{k'}k
%     % }{
%     %   \E \vdash \Cleq{\forall \kvar_i \forall (\tvar_j:k_j).\ \qual{C_x}{\tau}}{k} \Crewrite  D
%     % }
%   \end{mathpar}
%   \caption{Auxiliary rules for inference}
%   \label{op:usgmap}
% \end{figure}

\input{infer/metatheory}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: