### 2019-11-18 kindly/infer/exampleconstr.tex
Put the constraint example in the body, and push all the constrain theory out.

\label{solving:example}

\begin{figure}[tbp]
  \centering
  \begin{tikzpicture}[node distance=7mm,every edge/.style=[link,->,>=latex,thick]]
    \node (U) {$\kun$} ;
    \node[above left=of U] (x) {$\kvar_x$} ;
    \node[above=of x] (r) {$\kvar_r$};
    \node[left=of x] (g) {$\kvar_\gamma$} ;
    \node[right=of x] (b) {$\kvar_\beta$} ;
    \node[right=of b] (3) {$\kvar_3$} ;
    \node[above=of 3] (f) {$\kvar_f$} ;
    \node[above=of f] (1) {$\kvar_1$} ;

    \draw (x) to[bend right] (U) ;
    \draw (x) -> (r) ;
    \draw (b) -> (r) ;
    \draw (g) -> (x) ;
    \draw (3) -> (f) ;
    \draw (f) -> (1) ;

    \draw[blue] (3) to[bend right] (1);

    \node at (-0.5,-0.5) {Before};

    \begin{scope}[dashed,gray]
      \draw (U) to[bend right] (x) ;
      \draw (U) to (b) ;
      \draw (U) to[bend left] (g) ;
      \draw (U) to (r) ;
      \draw (U) to (1) ;
      \draw (U) to (3) ;
      \draw (U) to (f) ;
    \end{scope}

    \begin{scope}[on background layer]
      \node[fill=green!20,draw=green,
      inner sep=-1pt,ellipse,rotate fit=-20,fit=(U.south) (x) (g)] {};
      \node[fill=red!20,draw=red, inner sep=-2pt, circle,fit=(r)] {};
      \node[fill=red!20,draw=red, inner sep=-2pt, circle,fit=(f)] {};
    \end{scope}
  \end{tikzpicture}~
  \vrule~
  \begin{tikzpicture}[node distance=7mm,every edge/.style=[link,->,>=latex,thick]]
    \node (b) {$\kvar_\beta$} ;
    \node[right=of b] (3) {$\kvar_3$} ;
    \node[above=of 3] (f) {} ;
    \node[above=of f] (1) {$\kvar_1$} ;

    \draw (3) to[bend right] (1);

    \node at (0.5,-1.5) {After};
  \end{tikzpicture}
    \caption{Graph representing the example constraints}
    \label{example:graph}
    % \caption{Final state of the graph}
    \label{example:graph:final}
\end{figure}

We consider the expression $\lam{f}{\lam{x}{(\app{f}{x},x)}}$.
The application of our inference algorithm yields the following constraints:
%
\begin{align*}
  \E &= (\tvar_f : \kvar_f)
  (\tvar_x : \kvar_x)\dots\\
  C &= (\tvar_f \leq \gamma \tarr{\kvar_1} \beta )
  \Cand
  (\gamma \leq \tvar_x)
  \Cand
  (\beta * \tvar_x \leq \alpha_r)
  \Cand
  (\kvar_x \leq \kun)
\end{align*}

The first step of the algorithm is to use Herbrand unification to obtain
a type skeleton. We thus obtain
$$\alpha_r =
(\gamma \tarr{\kvar_3} \beta) \tarr{\kvar_2} \gamma \tarr{\kvar_1} \beta * \gamma$$

and the following kind constraints: 
\[\begin{aligned}
    % \Cproj{\kvar_r}\Cproj{\kvar_f}\Cproj{\kvar_x}{}
    &(\kvar_x \leq \kun)
    \Cand
    (\kvar_\gamma \leq \kvar_x)
    \Cand
    (\kvar_x \leq \kvar_r)
    \Cand\\
    &(\kvar_\beta \leq \kvar_r)
    \Cand
    (\kvar_3 \leq \kvar_f)
    \Cand
    (\kvar_f \leq \kvar_1)
\end{aligned}\]
% Since $\kvar_r$, $\kvar_f$ and $\kvar_x$ are not present in the type skeleton,
% they are existentially quantified.

We translate these constraints into a relation whose graph
is shown in \cref{example:graph}.
%
The algorithm then proceeds as follow:
\begin{itemize}[noitemsep]
\item From the constraints above, we deduce the graph shown
  with plain arrows on the left
\item We add all the dashed arrows by saturating
  lattice inequalities. For clarity, we only show $\kun$.
\item We identify the connected component circled in
  {\color{green} green}.
  We deduce $\kvar_\gamma = \kvar_x = \kun$
\item We take the transitive closure, which adds the
  arrow in {\color{blue} blue} from $\kvar_1$ to $\kvar_3$.
\item We remove the remaining nodes not present in the type skeleton (colored in {\color{red} red}): $\kvar_r$ and $\kvar_f$.
\item We clean up the graph (transitive reduction, remove unneeded constants, \dots),
  and obtain the graph shown on the right.
  We deduce $\kvar_3 \leq \kvar_1$.
\end{itemize}

The final constraint is thus
$$\kvar_\gamma = \kvar_x = \kun \Cand \kvar_3 \leq \kvar_1$$
If we were to generalize, we would obtain the type scheme:
$$\forall \kvar_\beta \kvar_1 \kvar_2 \kvar_3
(\gamma : \kun) (\beta : \kvar_\beta).\ %
\qual
{\Cleq{\kvar_3}{\kvar_1}}
{(\gamma \tarr{\kvar_3} \beta) \tarr{\kvar_2} \gamma \tarr{\kvar_1} \beta * \gamma}$$

The algorithm we presented does not use variance, however we can use it to
simplify this type further by noticing that $\kvar_1$ and $\kvar_2$ are only
used in covariant position, and can thus be replaced by their lower bound.
By removing the unneeded quantifier, we obtain the equivalent type:
$$
\forall \kvar
(\gamma : \kun).
{(\gamma \tarr{\kvar} \beta) \tarr{} \gamma \tarr{\kvar} \beta * \gamma}$$

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: