### 2019-11-22 kindly/infer/infer.tex
Remove pairs from the body.

\section{Inference}
\label{inference}

An important contribution of \affe is its principal type inference.
We now formulate our inference algorithm
based on the $HM(X)$ framework~\citep{DBLP:journals/tapos/OderskySW99},
a Hindley-Milner type system for a language
with qualified types where constraints are expressed in an arbitrary
theory $X$.
Most importantly,
$HM(X)$ ensures principal type inference
when $X$ respects the specified properties.
We apply HM(X) to a concrete constraint language which we name $\CL$.
We adapt and extend HM(X)'s existing rules to support kind inference,
track linearity and handle borrows and regions. Finally, we
formulate an appropriate constraint solving algorithm for $\CL$
which solves and simplifies constraints in a principled way.

\subsection{Preliminaries}

In the context of inference, it's critical to know which elements
are input and output of inference judgements.
In the following, when presenting a new judgement,
we write input parameters in \textbf{\textcolor{DarkSlateBlue}{bold blue}}. The rest will be
output parameters.

\paragraph{Usage environments}

% One novelty of our inference  judgement now
% returns a type environment, which we call ``usage environment'' and commonly
% write $\Sv$, to summarize how variables and borrows are used inside
% the expression.

To determine if a variable is used in an affine manner, we must track
its uses and the associated kinds. In the expression
$f\ x\ x$, $x$ is used twice. If $x$ is of type $\tau$, which is of kind $k$,
we add the constraint $\Cleq{k}{\kun}$.
%
To infer such constraints, our inference judgement not only
take an environment as parameter but also return an environment called
``usage environment'', denoted $\Sv$,
which summarizes usages of variables and borrows.
Usage environments follow the grammar
of normal environments.
%
In \cref{sdtyping}, we used relations to split environments in two and to
transform suspended bindings into borrows inside a region.
These relations take as argument a constraint which validates
the transformations.
In the context of inference, we define new judgements which \emph{infer}
the constraints.
\begin{itemize}[leftmargin=*]
\item $\bsplit{C}{\Sv}{\inP{\Sv_1}}{\inP{\Sv_2}}$.
  Given two usage environments $\inP{\Sv_1}$ and $\inP{\Sv_2}$,
  we return $\Sv$, the merged environment, and $C$, a set
  of constraints that must be respected.
\item $\bregion[\inP{n}]{C}{\inP{x}}{\Sv}{\inP\Sv'}$.
  Given a usage environment $\inP{\Sv'}$, a nesting level $\inP{n}$,
  and a variable name $\inP{x}$, we return
  $\Sv$ where the borrow binding of $x$ in $\Sv'$, if it exists,
  is replaced by
  a suspended binding. We also return the constraints $C$.
\end{itemize}
Both relations are total and non-ambiguous in term of their input
(i.e., functions), and use
the rules presented in \cref{sdtyping:split,sdtyping:regions}.
%
The relations used for syntax-directed typing can trivially be defined
in terms of these new relations by using constraint entailment.
All relations are fully described in \cref{typ:extra:envs}.

\paragraph{Constraint normalization}

The HM(X) framework assumes the existence of a ``$\operatorname{normalize}$''
function which takes a constraint $\inP{C}$, a substitution $\inP\psi$ and returns a
simplified constraint $C'$
and an updated substitution $\unif'$.
Normalization returns a normal form such that $\unif'$ is a most general unifier.
For now, we simply
assume the existence of such a function for our constraint system
and delay details to \cref{infer:solving}.

\subsection{Type inference}

We write $\inferW{\Sv}{(C,\unif)}{\inP{\E}}{\inP{e}}{\tau}$ when
$\inP{e}$\ has type $\tau$ in $\inP{\E}$ under the constraints $C$ and unifier $\unif$
with a usage environment $\Sv$. $\inP\E$ and $\inP{e}$ are the input parameters of our
inference algorithm.
Unlike in the syntax-directed version, $\E$ contains only regular and type bindings.
Suspended and borrow bindings can only be present in $\Sv$.
We revisit some of the syntax-directed rules presented in \cref{sdtyping}
to highlight the novelties
of our inference algorithm and the differences with the syntax-directed
system in \cref{rule:infer:envs}.
The complete type inference rules are shown in \cref{appendix:infer}.

\paragraph{Environments and bindings}
\label{infer:envs}
%
\begin{figure*}[tb]
  \begin{mathpar}
    \ruleIVar
    \and
    \ruleIAbs
    \and
    % \ruleILet
    % \hfill
    \ruleIRegion
    % \and
    % \ruleIPair
    \and
    \ruleIApp
    \and
    \inferrule{}
    { \Weaken_{\bvar{x}{\sigma}}(\Sv) =
      \text{if } (x\in\Sv) \text{ then} \Ctrue \text{else }
      \Cleq{\sigma}{\kaff_\infty}
      % \begin{dcases}
      %   \Ctrue&\text{if } x\in\Sv\\
      %   \Cleq{\sigma}{\kaff_\infty}&\\
      % \end{dcases}
      % \\\\\\
      % \generalize{C}{\E}{\tau} =
      % (\Cproj{(\Multi{\kvar},\Multi{\tvar})}{C},
      % \forall (\Multi{\kvar},\Multi{\tvar}).\qual{C}{\tau})\\
      % \text{ where}\\\\
      % \Multi{\kvar},\Multi{\tvar} = (\fv{\tau}\cup\fv{C})\setminus\fv{\E}
    }
  \end{mathpar}
  \vspace{-15pt}
  \caption{Selected inference rules -- $\inferW{\Sv}{(C,\unif)}{\inP{\E}}{\inP{e}}{\tau}$}
  \label{rule:infer:envs}
  \label{rule:infer:envrules}
  \label{rule:infer:let}
\end{figure*}
%
In the syntax-directed system, the {\sc Var} rule ensure
that linear variables are not discarded at the \emph{leaves}.
In the inference algorithm, we operate in the opposite
direction: we collect data from the leaves and enforce linearity
at \emph{binders}. This is reflected in the {\sc Var$_I$} and
{\sc Abs$_I$} rules.
Typing for variables is very similar to traditional (non-affine) HM.
The difference is that we record
that $x$ was used with the scheme $\schm$ by returning
a usage environment $\Sv = \{ \bvar{x}{\schm} \}$.
%
This usage environment is in turn used at each binder to enforce proper
usage of linear variable via the $\Weaken$ property.
We demonstrate this for lambda expressions
in the {\sc Abs$_I$} rule.
First, we typecheck the body of the lambda and obtain a usage
environment $\Sv_x$. Like in the syntax-directed type system,
we introduce the constraint
$\Cleq{\Sv\Sdel{x}}{\kvar}$ which properly accounts for captures in
the body of the lambda expression. We then introduce the constraint
$\Weaken_{\bvar{x}{\sigma}}(\Sv)$, which is false only if weakening
is necessary-yet-impossible due to linearity.
The $\Weaken$ constraint is introduced at each binding point in the program.
%
Finally, we normalize constraints to ensure
that the inference algorithm always return the simplest possible
constraints and unifiers.

\paragraph{Splitting and Regions}
\label{infer:split}
\label{infer:regions}

Inference versions
of the {\sc App} and {\sc Region} rules
are similar to the original ones, but now \emph{return} the usage
environment $\Sv$.
As such, we use the ``inference'' version of the relations on
the environment,
$\bsplit{C}{\Sv}{\inP{\Sv_1}}{\inP{\Sv_2}}$
and $\bregion{C}{\inP{x}}{\Sv}{\inP\Sv'}$,
which returns the necessary constraints.
We then collect all constraints and normalize them.

\subsection{Constraint solving}
\label{infer:solving}

\newcommand\A{\mathcal A}
\newcommand\SC{\mathcal S}

%
% \begin{figure}[tb]
%   \centering
%   \begin{align*}
%     C &::= \Cleq{\tau_1}{\tau_2}
%         \mid \Cleq{k_1}{k_2}
%         \mid C_1 \Cand C_2
%         \mid \Cproj{\tvar}{C}
%         \mid \Cproj{\kvar}{C}
%   \end{align*}
%   \caption{The constraint language}
%   \label{grammar:constraint}
%   % \input{infer/lattice}
%   % \caption{Lattice inequalities -- $k \lk_\Lat k'$}
%   \input{infer/entails}
%   \caption{Base entailment rules -- $\entail{\inP{C}}{\inP{D}}$ }
%   \label{rules:entail}
% \end{figure}

For concision, we demonstrate our constraint solving
algorithm on an example. The complete constraint system is
defined in \cref{appendix:constraints}.
%
\input{infer/exampleconstr}

Our algorithm respect the properties
of HM(X), computes principal normal forms
and simplifies constraints significantly.
All the simplification mechanisms presented
here, including the variance-based one, are complete.
It is also possible to add ``best-effort'' simplification
rules which help reduce the size of inferred signatures even further
\citep{DBLP:conf/aplas/Simonet03}.

% 
% We now show that this algorithm is well-behaved with respect to entailment
% and computes unique normal forms.

% \begin{property}[Principal normal form]
%   Normalization computes principal normal forms for $\CL$, i.e.
%   given a constraint $D\in\CL$, a substitution $\phi$ and
%   $(C,\unif) = \normalize{D}{\phi}$,
%   then $\phi\leq\unif$,
%   $C \equivC \unif D$ and
%   $\unif C = C$.
% \end{property}

% \begin{property}[Principal constraints]
%   $\CL$ has the principal constraint property, i.e.
%   for every $D\in\CL$, and a substitution $\unif$,
%   either $D$ does not have a normal form, or it has
%   a principal normal form.
% \end{property}

% \begin{property}[Regular constraint system]
%   $\CL$ is regular, ie, for $x, x'$ two types or kinds,
%   $\entail{}{\Ceq{x}{x'}}$ implies
%   $\fv{x} = \fv{x'}$
% \end{property}

% The proofs are given in \cref{appendix:constraints}. These three properties
% are sufficient to state that $HM(\CL)$ provides principal type inference.
% However, we use an extension of HM(X) with kind inference
% and affine types and borrows.
% We now show that HM(X)'s completeness and soundness theorems are still valid on
% this extended system.

\subsection{Soundness and Principality}

We now state that our extended inference algorithm still satisfies the soundness
and completeness properties of HM(X).
% This is achieved by extending
% the original proofs from HM(X), which is done in \cref{appendix:infer}.
%
The first theorem states that our inference algorithm is sound
with respect to the syntax-directed type system.

\begin{theorem}[Soundness of inference]
  Given a type environment $\E$ containing only value bindings,
  $\E|_\tau$ containing only type bindings, and a term $e$:\\
  if $\inferW{\Sv}{(C,\unif)}{\E;\E_\tau}{e}{\tau}$\\
  then $\inferS{C}{\unif(\Sv;\E_\tau)}{e}{\tau}$, $\unif C = C$ and $\unif \tau = \tau$
\end{theorem}

The syntax-directed derivation holds with the usage environment $\Sv$ instead of the originally provided environment $\E$. Indeed,
$\E$ does not contain suspended and borrow bindings, since those
are discovered on the fly and recorded in $\Sv$. The type binding, however,
are directly taken from the syntax-directed derivation.

The second theorem states that our algorithm is complete: for any given
syntax-directed typing derivation, our inference algorithm can find
a derivation that gives a type at least as general.
For this, we need first to provide a few additional definitions.

% \begin{definition}[More general unifier]
%   Given a set of variable $U$ and $\unif$, $\unif'$ and $\phi$
%   substitutions. \\
%   Then
%   $\unif \leq^{\phi}_{U} \unif'$ iff $(\phi \circ \unif)|_{U} = \unif'|_U$.
% \end{definition}

\begin{definition}[Instance relation]
  Given a constraint $C$ and two schemes
  $\schm = \forall \Multi\tvar. \qual{D}{\tau}$ and
  $\schm' = \forall \Multi\tvar'. \qual{D'}{\tau'} $.
  Then $\entail{C}{\schm \preceq \schm'}$
  iff $\entail{C}{\subst{\tvar}{\tau''} D}$
  and $\entail{C\Cand D'}{\Cleq{\subst{\tvar}{\tau''}\tau}{\tau'}}$
\end{definition}

\begin{definition}[Flattened Environment]
A flattened environment,
written as $\Eflat\E$, is the environment
where all the binders are replaced by normal ones. More formally:
\begin{align*}
  \Eflat\E
  =& \left\{\bvar{x}{\tau}\in\E \mid
    \vee \bvar{\borrow{x}}{\borrowty{k}{\tau}}\in\E
    \vee \svar{x}{\tau}^n\in\E
    \right\}\\
  &\cup \left\{ \bvar{\tvar}{k} \mid \bvar{\tvar}{k}\in\E \right\}
\end{align*}
\end{definition}

We can then state principality.

% \begin{theorem}[Completeness]
%   Given $\inferS{C'}{\E'}{e}{\tau'}$ and
%   $\entail{}{\unif'\E \preceq \E'}$.
%   Then $$\inferW{\Sv}{(C,\unif)}{\Eflat\E}{e}{\tau}$$
%   for some environment $\Sv$,
%   substitution $\unif$, constraint $C$ and type $\tau$ such
%   that
%   \begin{align*}
%     \unif &\leq^{\phi}_{\fv{\E}} \unif'
%     &\entail{C'&}{\phi C}
%     &\entail{&}{\phi \schm \preceq \schm'}
%     &\Sv&\subset\E
%     % &( C, \schm, \unif) &\leq (C',\schm',\unif')
%   \end{align*}
%   where $\schm' = \generalize{C'}{\E'}{\tau'}$
%   and $\schm = \generalize{C}{\E}{\tau}$
% \end{theorem}

% Finally, principality is a direct application of completeness to
% toplevel programs.

\begin{theorem}[Principality]
  Let $\inferS{\Ctrue}{\E}{e}{\schm}$ a closed typing judgement.\\
  Then $\inferW{\Sv}{(C,\unif)}{\Eflat\E}{e}{\tau}$
  such that:
  \begin{align*}
    (\Ctrue,\schm_o) &= \generalize{C}{\unif\E}{\tau}
    &\entail{&}{\schm_o \preceq \schm}
    % &( C, \schm, \unif) &\leq (C',\schm',\unif')
  \end{align*}

\end{theorem}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: