### 2020-02-28 kindly/main.tex
abstract

\PassOptionsToPackage{svgnames}{xcolor}
\documentclass[acmsmall,review,anonymous]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

% Encoding and lang
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

% Graphical packages
\usepackage{graphicx}

\usepackage{xspace}

% Math
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{thm-restate}
% \usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{textcomp}
\usepackage{gensymb}
% \usepackage{textgreek}
\usepackage{multicol}

\usepackage{listings}
\usepackage[scaled=0.85]{DejaVuSansMono}

\usepackage{tikz}
\usetikzlibrary{arrows}

\usepackage{caption}
\usepackage{subcaption}
\usepackage[inline,shortlabels]{enumitem}
\setlist{leftmargin=*,noitemsep}
\usepackage{array}
\usepackage{bm}%Decent bolding for math symbols

\usepackage{natbib}% Good citations and bibliography
\usepackage{mathpartir} % Syntax trees

\mprset{sep=1.5em}

\usepackage{tikz}
\usetikzlibrary{decorations.text,backgrounds,positioning,shapes,
  shadings,shadows,arrows,decorations.markings,calc,fit,fadings,
  tikzmark
}

\usepackage[nameinlink,capitalize]{cleveref}
\addto\extrasenglish{%
  \def\sectionautorefname{Section}
  \def\subsectionautorefname{Section}
  \def\subsubsectionautorefname{Section}
  \def\lstnumberautorefname{Line}
}
\newtheorem{property}{Property}
%\newtheorem{corollary}{Corollary}

\bibliographystyle{ACM-Reference-Format}

\author{Gabriel Radanne}
\affiliation{
  \institution{University of Freiburg}
  \country{Germany}
}
\email{radanne@informatik.uni-freiburg.de}

\author{Hannes Saffrich}
\affiliation{
  \institution{University of Freiburg}
  \country{Germany}
}
\email{saffrich@informatik.uni-freiburg.de}

\author{Peter Thiemann}
\affiliation{
  \institution{University of Freiburg}
  \country{Germany}
}
\email{thiemann@informatik.uni-freiburg.de}

\input{prelude}
\input{caml}
\input{lst-rule}
\input{notations}

\begin{document}

\title{Kindly Bent to Free Us}
\begin{abstract}
  Systems programming often requires the manipulation of resources like
  file handles, network connections, or dynamically allocated
  memory. Programmers need to follow certain protocols to handle
  these resources correctly. Violating these protocols causes bugs
  ranging from type mismatches over data races to use-after-free
  errors and memory leaks. These bugs often lead to security  vulnerabilities.

  While statically typed programming languages guarantee type soundness and memory
  safety by design, most of them do not address issues arising
  from improper handling of resources.
  An important step towards handling resources is the adoption of 
  linear and affine types that enforce single-threaded resource usage.
  However, the few languages supporting such types require heavy type
  annotations.

  We present \lang, an extension of ML that
  manages linearity and affinity properties using kinds and
  constrained types. In addition
  \lang{} supports the exclusive and shared borrowing of affine
  resources, inspired by features of Rust.
  Moreover, \lang{} retains the defining features of the ML family:
  it is an impure, strict, functional expression language with complete principal type
  inference and type abstraction.
  \lang{} does not require any linearity annotations in
  expressions and supports common functional programming idioms.
\end{abstract}

\maketitle

\input{rulemacros}

\input{intro}
\input{motivation}
\input{lang}
\input{infer/infer}
\input{metatheory}
\input{soa}
\input{conclusion}

\bibliography{biblio}

\clearpage
\appendix

\section{Further Examples}
\label{sec:extra-example}
\lstMakeShortInline[keepspaces,basicstyle=\small\ttfamily]@
\input{motivation-linearity}
\input{motivation-pool}
\lstDeleteShortInline@

\input{sdtyping/extra}
\clearpage
\input{infer/extra}
\clearpage
\input{lang-extra}
\input{metatheory-extra}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: