### 2020-03-04 kindly/extensions.tex
Small additions.

\lstMakeShortInline[keepspaces,basicstyle=\small\ttfamily]@
\section{Extensions}

\subsection{Algebraic Datatypes}

Algebraic data types are a staple of functional programming and fit nicely
in our paradigm. Indeed, it is sufficient to ensure  that the kinds of
the constructor arguments are less than or equal to the kind of the datatype.
Hence, it is forbidden to include affine elements in an unrestricted
datatype, whereas the elements in a linear list may be linear or
unrestricted. 
Here is the definition of a datatype for linear lists.

\begin{lstlisting}
type ('a : 'k) llist : lin = Nil | Cons of ('a * 'a llist)
\end{lstlisting}

This extension is implemented in our prototype.

\subsection{Ad-hoc Polymorphism and Borrows}

In our formalization, we use two operators, $\borrow{x}$ and $\reborrow{x}$ to
distinguish between borrows and borrows of borrows.
Such a distinction is inconvenient for programming.
Using a typeclass-like mechanism, we can replace these operators
by a single overloaded operator, $\borrow{x}$, which expects $x$ to be @Borrowable@ and
would then desugar to the more precise operators.
A similar solution is used in Rust through the @Borrow@ and @Defer@
traits.
This approach also enables method calls on objects without
explicit borrows, such as @foo.len()@ where len expects a shared borrow.

Ad-hoc polymorphism fits demonstrably in the \hmx framework of constrained
types and preserves all properties of our language such
as principal type inference. Its soundness is orthogonal to linear types
and has been explored in the literature~\citep{DBLP:conf/fpca/OderskyWW95}.

\subsection{Branching constructs}

Our formalization of \lang does not contain branching constructs. As a consequence,
subexpressions are always executed sequentially. This is of great importance
for the splitting relation introduced in \cref{sdtyping}, as it must
coincide with the order of evaluation. To account for borrows and linear
values in different branches, we need a new (symmetric) join relation.
This relation immediately follows from the semantics of borrows and linearity
shown so far, and allow to easily define branching constructs such as if-then-else
and pattern matching.
This extension is implemented in our prototype.

\subsection{Non-Lexical Regions}

A recent important addition to Rust is the notion of non-lexical lifetimes.
With this feature code is acceptable even if borrowing does not respect
lexical scoping as in this example:

\begin{lstlisting}
let a = &x in
f a;
g (&!x)
\end{lstlisting}

Such code patterns are dynamically safe since $a$ is not used after
the second line.
Non-lexical lifetimes handle this by removing expressions
that do not mention $a$ from its region; in this example,  the
expression on the last line.
%
In \lang, regions are purely lexical and marked by the
expression $\region{b}{e}$.
During inference, we introduce kind constraint to prevent escaping from
a region.

We can replace the lexical region by
an annotation on each expression indicating which borrows are live
in this expression.
When exploring a subexpression, we compare the annotations, and automatically
apply the $\textsc{Region}$ rule when they differ.
%
This approach is equivalent to inlining the $\textsc{Region}$ rule in all the other
rules.

Applied to the program above, only the first two lines would be annotated
to be ``in the region associated with @&x@'', but not the last line.
Thanks to these annotations, when type checking the sequence we check
that the borrow does not escape the left-hand side (i.e., the second line).

\lstDeleteShortInline@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: