### 2020-04-23 kindly/icfp20-rebuttal.md
First version of the ICFP rebuttal.

We thank the reviewers for their thoughtful comments.
We will take their comments into account and revise our article as follows.
- Expand and improve our examples
- Precise further the limitations of our system and the comparison to other systems

# Reviewer A

> The use of "absolute" natural numbers to identify regions seems not modular.

Affe requires that each region is identified by an index along with a partial
order on indices that is compatible with the nesting of regions.
This can be implemented in many ways, including region variables
often used in algebraic effects systems, existentials, etc.

Natural numbers are the absolute simplest way to implement such indices
which is why
we use them in the formalisation. Our formalisation only relies on
the partial order and could easily use any of these
more sophisticated systems. This would certainly be the case in a 
realistic implementation for OCaml, for instance, to leverage
the on-going work on algebraic effects.
The properties of the system would stay identical.

We plan to add this point to section 6.

> It is very hard to get a grasp what kind of programming idioms can be type checked using Affe.

Affe can easily express rich data-structures (notably 
persistent ones) or manipulate resources and protocols (for instance with session types).

The main limitations of Affe is that it is not flow sensitive.
Code idioms that relies of subtle flow-sensitive usage of 
permissions and linearity will not work in Affe. Such patterns often
do not work in Rust either, and require richer logic more akin to theorem
proving, such as Mezzo.

Rust slightly relaxes these constraints through the use of non-lexical lifetimes
and double-borrows. 
Following the work by Weiss et al. [38] on the Rust semantics, we believe
these techniques can also be applied to Affe and 
sketch how in section 6.

> What is the reason the borrow is fixed to `U` in the splitting rules `SuspB` and `SuspS` (page 10).

Peter/Hannes ? Is this due to a proof thingy ? In principle, doing an exclusive reborrow is ok.

> Is Affe's kind mechanism also powerful enough to keep track of things like `Send` and `Sync`?

This is probably better implemented via ad-hoc polymorphism. It is very
desirable to extend Affe with some form of typeclasses, as we point out in
Section 6.2, which would allow to implement such mechanisms.

# Reviewer B

> What could go wrong, for example, if the set function of the array module would use a non-exclusive borrow type in the argument?

This question should be reframed in a larger context: "What does my API guarantee ?".
For instance, ML references and arrays allow concurrent writes in a perfectly
safe manner. In a more general concurrent context, 
many data-structures are designed to allow concurrent mutations, via mutexes for
instance.

However, if the API prevents concurrent mutations, this ensures that no data-races
can happen on this data-structure. Data-races are often the source of subtle
and hard-to-fix bugs.
Exclusive borrow allows the author to choose if they want or not to provide
such guarantee. This has been used with huge success in the Rust community to 
write safe and fast concurrent programs.

# Reviewer C

> in other systems, one might use existential quantification tricks or magic wands to express temporarily borrowing an *element* of an array before returning the element to the array. What would be the equivalent here?

Indeed, this is something that we would like to add to the system. 
Such functions can already be implemented via closures, 
as demonstrated in the iteration examples in section 2.3:

    val get_borrow :
      ('b <= affₖ) < &('a t) * int * (&(linₖ₊₁, 'a) -{linₖ₊₁}> 'b) -> 'b

This API is however a bit inconvenient. We can improve it using
so called "existential tricks" to allow matching on borrows:

    let &x = get_borrow (a,3) in (* &x must not escape here *)

In general, existentials tricks are not incompatible with Affe and are in fact
one of the foremost way to extend its expressivity.

> Do you envision a similar discipline for your system where one might write, say, a doubly-linked list in regular GC'd ML, then assume an interface for it that exposes it as a resource?

This pattern is indeed essential and was well demonstrated by Alms.
Affe already allow such patterns via subkinding and module type abstraction:
a type can be defined as unrestricted and exposed as linear.

Integrating "forced freeing" of such objects with the GC is more delicate
and is explored by Munch-Maccagnoni [21].