### 2022-12-07 kindly/slides-icfp.tex
slides v2022

\documentclass[aspectratio=169,dvipsnames,svgnames,10pt]{beamer}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%\usepackage[french, english]{babel}
%\selectlanguage{english}

\usepackage{graphicx}

% Math
\usepackage{amsmath}
% \usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
% \usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{textcomp}
% \usepackage{textgreek}

% Specialized packages
% \usepackage{syntax} % Grammar definitions
\usepackage{verbatim}
\usepackage{listings} % Code
\usepackage{xspace} % Useful for macros
%\usepackage{natbib}% Good citations and bibliography
%\usepackage{mathpartir} % Syntax trees
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{multicol}%multicolumn lists
\usepackage{pifont}%% check mark
\usepackage{minted}
\usepackage{wasysym}

% \usepackage{mathptmx}
% \usepackage[scaled=0.9]{helvet}
\usepackage{beramono}

% \bibliographystyle{ACM-Reference-Format}
% \citestyle{acmauthoryear}
\usepackage{appendixnumberbeamer}

\usetheme{metropolis}
\beamertemplatenavigationsymbolsempty
\setbeamercovered{transparent=20}
\metroset{
  sectionpage=none,
  numbering=fraction,
  progressbar=foot,
}

\usepackage{tikz}
\usetikzlibrary{decorations.text,backgrounds,positioning,shapes,
  shadings,shadows,arrows,decorations.markings,calc,fit,fadings,
  tikzmark,scopes
}
\input{tikz}

\def\HUGE{\fontsize{35pt}{15pt}\selectfont}

\input{prelude}
\input{caml}
\input{notations}
\input{rulemacros}

\newcommand\Y{{\color{Green}{\ding{52}}}}
\newcommand\N{{\color{Red}{\ding{56}}}}
\newcommand\M{{\color{Orange}{\textasciitilde}}}

\lstset{
  tabsize=4,
  aboveskip={0.5\baselineskip},
  belowcaptionskip=0.5\baselineskip,
  basicstyle=\scriptsize\ttfamily,
}

\title{Kindly Bent To Free Us}
\author{{Gabriel {Radanne}}
  \and Hannes {Saffrich}
  \and \textbf{Peter {Thiemann}}}
\date{ICFP 2020}

\begin{document}

\lstMakeShortInline[keepspaces,basicstyle=\small\ttfamily]@

\frame[plain]{\titlepage}

\begin{frame}
  \frametitle{\og high severity security bugs\fg in Chromium}

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.7\textwidth]
    {chromium-use-after-free}
  \end{figure}

  \visible<2->{
    Chromium is written in C/C++!\\
    Surely these bugs don't happen in high-level typed languages%
    \alt<3>{ \dots right?}{.}
  }

\end{frame}

\begin{frame}[fragile]
  \frametitle{Let's write some OCaml code}

\begin{minted}{OCaml}
let gradeKV : kv_store = KVs.opendbm "grades" ... in
KVs.add gradeKV "math" 42;
(* ... *)
KVs.close gradeKV;
(* ... *)
print_int (KVs.find gradeKV "literature") (* run-time error! *)
\end{minted}

\end{frame}

\begin{frame}[fragile]
  \frametitle{Affe!}

  \begin{onlyenv}<1-3>
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let gradeKV : #\tikzcoord{db1}#kv_store#\tikzcoord{db2}# = KVs.opendbm "grades" ... in
KVs.add #\tikzcoord{A1}#&!gradeKV#\tikzcoord{A2}# #\tikzcoord{string1}#"math"#\tikzcoord{string2}# 42;
(* ... *)
KVs.close gradeKV;
(* ... *)
print_int (KVs.find #\tikzcoord{U1}#&gradeKV#\tikzcoord{U2}# "literature") (* \N compile-time error! *)
\end{minted}
  \end{onlyenv}
  \begin{onlyenv}<4>
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let gradeKV #\tikzcoord{eq1}#=#\tikzcoord{eq2}# KVs.opendbm "grades" ... in
KVs.add &!gradeKV "math" 42;
(* ... *)
KVs.close gradeKV;
(* ... *)
print_int (KVs.find &gradeKV "literature") (* \N compile-time error! *)
\end{minted}
  \end{onlyenv}

  \begin{tikzpicture}[remember picture,overlay]
    \begin{onlyenv}<2>
      \node[code,fit=(db1) (db2)] (db) {};
      \node[code,fit=(string1) (string2)] (string) {};
      \node[overlay% ,text width=8.7cm
      ,annot,scale=1.2] at (5,-0.7) {
        Kinds restrict usage
        % \begin{itemize}
        % \item Linear (@lin@): Used exactly once $[1]$
        % \item Affine (@aff@): Used at most once $[0-1]$
        % \item Unrestricted (@un@): Used arbitrarily many time $[0-\infty]$
        % \end{itemize}
      };
      \draw[overlay,link] (db) to[out=60,in=-90]
      +(0,1) node[annot,anchor=south]
      {@type kv_store : lin (* Linear kind *)@};
      \draw[overlay,link] (string) to[out=-30,in=180]
      +(1.5,-0.5) node[annot,anchor=west]
      {@type string : un (* Unrestricted kind *)@};
    \end{onlyenv}
    \begin{onlyenv}<3>
      \node[code,fit=(A1) (A2)] (A) {};
      \node[code,fit=(U1) (U2)] (U) {};
      \node[annot,scale=1.2] at ($(A)+(2,-0.8)$) (borrows) {Borrows!};
      \draw[overlay,link,<-] (A) to[out=-30,in=150] (borrows) ;
      \draw[overlay,link,<-] (U) to[out=120,in=-70] (borrows) ;
    \end{onlyenv}
    \begin{onlyenv}<4>
      \node[code,fit=(eq1) (eq2)] (eq) {};
      \draw[overlay,link,<-] (eq) to[out=90,in=-90]
      +(0.5,1) node[annot,scale=1.2,anchor=south]
      {Complete Type Inference};
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}

\begin{frame}{Table of contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections]
\end{frame}

\section{Linearity through kinds}

\begin{frame}[fragile]
  \frametitle{Linearity through kinds}

  \textbf{Kinds restrict usage:}
  \begin{itemize}
  \item Linear (@lin@): Used exactly once $[1]$
  \item Affine (@aff@): Used at most once $[0-1]$
  \item Unrestricted (@un@): Used arbitrarily many time $[0-\infty]$
  \item<2-> Kinds can be polymorphic!
  \end{itemize}

  \textbf{Examples:}
  \begin{onlyenv}<1>
\begin{minted}[texcomments=true]{OCaml}
type kv_store : lin
type string : un
\end{minted}
  \end{onlyenv}
  \begin{onlyenv}<2>
\begin{minted}[texcomments=true]{OCaml}
type ('a : 'k) list : 'k
\end{minted}
  \end{onlyenv}
  \begin{onlyenv}<3->
\begin{minted}[texcomments=true]{OCaml}
type ('a : 'k) list : 'k
val create_list : ('a : un) => int -> 'a -> 'a list
\end{minted}

    \begin{onlyenv}<4>
      We also use \textbf{subkinding}: @un <= aff <= lin@
    \end{onlyenv}
  \end{onlyenv}
\end{frame}

\section{Functions and captures}

\begin{frame}{Table of contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections,currentsection]
\end{frame}

\begin{frame}[t,fragile]
  \frametitle{Functions and captures}

\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let gradeKV = KVs.open ...

let #\tikzcoord{fun1}#log_n_close#\tikzcoord{fun2}# msg = 
  printf "Closing: %s" msg;
  KVs.close #\tikzcoord{db1}#gradeKV#\tikzcoord{db2}#
\end{minted}
  \begin{onlyenv}<3->
    We infer the type:
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
val log_n_close : string #\tikzcoord{arr1}##$\tarr{\texttt{lin}}$##\tikzcoord{arr2}# unit
\end{minted}
  \end{onlyenv}

  \begin{tikzpicture}[remember picture,overlay]
    \begin{onlyenv}<2->
      \node[code,fit=(fun1) (fun2)] (fun) {};
      \node[code,fit=(db1) (db2)] (db) {};
      \node[annot,scale=1.2]
      at ($(fun)+(5,0.2)$) (capture) {
        Capture!
      };
      \draw[overlay,link] (fun) to[out=20,in=170] (capture) ;
      \draw[overlay,link] (capture) to[out=-90,in=0] (db) ;
    \end{onlyenv}
    \begin{onlyenv}<3->
      \node[code,minimum height=18pt,fit=(arr1) (arr2)] (arr) {};
      \draw[overlay,link,<-] (arr) to[out=-110,in=70]
      +(-1,-1) node[annot,scale=1.2,anchor=north] {
        Usage mode
      };
    \end{onlyenv}
    \begin{onlyenv}<4>
      \node[annot,scale=1.2,red] at ($(arr)+(4,1)$) (warn) {
        \textbf{Warning}: Does not say anything about the arguments
      };
      \draw[overlay,link,<-,red] (arr.north) to[out=70,in=-150] (warn.-175) ;
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}

\section{Borrows and regions}

\begin{frame}{Table of contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections,currentsection]
\end{frame}

\begin{frame}[t,fragile]{Borrows}

  Let @a@ be a (linear or affine) resource. 
  A \emph{borrow} is a temporary loan of @a@.
  \begin{itemize}
  \item
    \textbf{Shared} borrows \tikzcoord{UX1}@&a@\tikzcoord{UX2} are for observing the resource
    % \textbf{Shared} borrows \tikzcoord{UX1}@&a@\tikzcoord{UX2} which are Unrestricted (@un@)
  \item
    \textbf{Exclusive} borrows \tikzcoord{AX1}@&!a@\tikzcoord{AX2} are for modifying the resource
    % \textbf{Exclusive} borrows \tikzcoord{AX1}@&!a@\tikzcoord{AX2} which are Affine (@aff@)
  \end{itemize}

  \begin{onlyenv}<2->
  A correct usage of borrows:
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let avg =
  (KVs.find #\tikzcoord{U1}#&gradeKV#\tikzcoord{U2}# "math" + KVs.find #\tikzcoord{U3}#&gradeKV#\tikzcoord{U4}# "compsci") / 2
  (* \Y Multiple shared borrows *)
in
KVs.add #\tikzcoord{A1}#&!gradeKV#\tikzcoord{A2}# "average" avg (* \Y One exclusive borrow *)
\end{minted}
  \end{onlyenv}  
  \begin{tikzpicture}[remember picture,overlay]
    \begin{onlyenv}<2->
      \node[code,color=yellow,fit=(A1) (A2)] (A) {};
      \node[code,color=red,fit=(U1) (U2)] (UA) {};
      \node[code,color=red,fit=(U3) (U4)] (UB) {};
      \node[annot] at ($(UA)+(3,1)$) (un) { Unrestricted -- @un@ } ;
      \draw[link] (un) to[out=180,in=80] (UA) ;
      \draw[link] (un) to[out=0,in=100] (UB) ;
      \node[annot] at ($(A)+(2,-1)$) (aff) { Affine -- @aff@ } ;
      \draw[link] (aff) to[out=180,in=-80] (A) ;
    \end{onlyenv}
    \node[code,color=yellow,fit=(AX1) (AX2)] {};
    \node[code,color=red,fit=(UX1) (UX2)] {};
  \end{tikzpicture}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Borrows -- Example of uses}

% \begin{onlyenv}<2>
% \begin{minted}[texcomments=true]{OCaml}
% let gradeKV = ... in
% let avg =
%  let grade topic = KVs.find &gradeKV topic in (* \Y Capture *)
%  (grade "math" + grade "compsci") / 2
% in
% KVs.add &!gradeKV "average" avg (* \Y Exclusive borrow *)
% \end{minted}
% \end{onlyenv}
\begin{onlyenv}<1>
  \textbf{Rule 1}: Cannot use a borrow and the resource itself simultaneously
\begin{minted}[texcomments=true]{OCaml}
let gradeKV = ... in
f (gradeKV, &gradeKV) (* \N Conflict: direct use and borrow! *)
\end{minted}
\end{onlyenv}%
\begin{onlyenv}<2>
  \textbf{Rule 2}: Cannot use an exclusive borrow and any other borrow simultaneously
\begin{minted}[texcomments=true]{OCaml}
let gradeKV = ... in
f (&!gradeKV, &gradeKV) (* \N Conflicting borrows! *)
\end{minted}
  Two shared borrows are ok
\begin{minted}[texcomments=true]{OCaml}
let gradeKV = ... in
f (&gradeKV, &gradeKV) (* \Y two shared borrows *)
\end{minted}
\end{onlyenv}%
\begin{onlyenv}<3->
  \textbf{Rule 3}: Borrows must not escape
  \begin{onlyenv}<3>
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let f () = 
  let gradeKV = ... in
  #\tikzcoord{R1}#let x = (&gradeKV, "mygrades") in#\tikzcoord{R2}#
  let clean_up = KVs.close gradeKV in
  #\tikzcoord{R3}#x#\tikzcoord{R4}#
  (* \N Borrow escaping its scope! *)
\end{minted}
  \end{onlyenv}
  \begin{onlyenv}<4->
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let f () = 
  let gradeKV = ... in
  #\tikzcoord{R1}#{| let x = (#\tikzcoord{B1}#&gradeKV#\tikzcoord{B2}#, "mygrades") in#\tikzcoord{R2}#
  let clean_up = KVs.close gradeKV in
  #\tikzcoord{R3}#x |}#\tikzcoord{R4}#
  (* \N Borrow escaping its scope! *)
\end{minted}
  \end{onlyenv}%
  \begin{tikzpicture}[remember picture,overlay]%
    \begin{onlyenv}<4->
      \node[code,rounded corners=0,color=red,fit=(R1) (R2) (R4)] {};
      \node[code,rounded corners=0,minimum height=15pt,yshift=1pt,color=red,fit=(R3) (R4)] {};
      \node[code,opacity=0,fit=(B1) (B2)] (B) {};
    \end{onlyenv}
    \begin{onlyenv}<5->
      \draw[link,<-] (B.10) to[in=180,out=40] ++(0.5,0.5) node[annot, anchor=west] {
        @&gradeDb : &(kv_store, un_2)@
      };
      \draw[link,<-] (R2) to[in=90,out=-10] ++(1.5,-0.5) node[annot, anchor=north] {
        Region nesting level: 1
      };
    \end{onlyenv}
  \end{tikzpicture}%
  \visible<5>{%
    \textbf{Indexed} kinds ensure that borrows do not escape!

    Borrows of index $2$ cannot escape a region of index $1$.
  }
\end{onlyenv}%
\end{frame}

% \begin{frame}[fragile]{API}

%   {\Large The \texttt{KVs} API:}
% \begin{minted}[texcomments=true]{OCaml}
% type kv_store : lin
% val find : &(kv_store,'k) -> string -{'k}> int
% val add : &!(kv_store,'k) -> string -{'k}> int -{'k}> unit
% \end{minted}

% \end{frame}

  % \begin{tikzpicture}[remember picture,overlay]
  %   \begin{onlyenv}<2->
  %     \node[code,color=yellow,fit=(A1) (A2)] {};
  %     \node[code,color=red,fit=(U1) (U2)] {};
  %     \node[code,color=red,fit=(U3) (U4)] {};
  %   \end{onlyenv}
  %   \node[code,color=yellow,fit=(AX1) (AX2)] {};
  %   \node[code,color=red,fit=(UX1) (UX2)] {};
  % \end{tikzpicture}

\begin{frame}[fragile]
  \frametitle{Everything together}

  The @Kv_Store@ API:
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
type kv_store : lin
val find : &(kv_store,'k) -> string #$\tarr{\texttt{'k}}$# int
val add : &!(kv_store,'k) -> string #$\tarr{\texttt{'k}}$# int #$\tarr{\texttt{'k}}$# unit
\end{minted}

  \begin{onlyenv}<2->
    A simple use:
  \end{onlyenv}
  \begin{onlyenv}<2>
\begin{minted}[texcomments=true]{OCaml}
let gradeKV = ... in
let avg =
  (KVs.find &gradeKV "math" + KVs.find &gradeKV "compsci") / 2

in
KVs.add &!gradeKV "average" avg
\end{minted}
  \end{onlyenv}
  \begin{onlyenv}<3>
\begin{minted}[texcomments=true]{OCaml}
let gradeKV = ... in
let avg =
  let grade subject = KVs.find &gradeKV subject in (* \Y Capture *)
  (grade "math" + grade "compsci") / 2
in
KVs.add &!gradeKV "average" avg
\end{minted}
  \end{onlyenv}
  \begin{onlyenv}<4>
\begin{minted}[texcomments=true]{OCaml}
let gradeKV = ... in
let avg =
  let grade = KVs.find &gradeKV in (* \Y Partial application *)
  (grade "math" + grade "compsci") / 2
in
KVs.add &!gradeKV "average" avg
\end{minted}
  \end{onlyenv}
  \begin{onlyenv}<5->
    \begin{onlyenv}<5>%
\begin{minted}[texcomments=true]{OCaml}
let average db subjects = 
  List.map (KVs.find db) subjects / List.length subjects
let main () =
  let gradeKV = ... in
  let avg = average &gradeKV ["math";"compsci";...] in
  KVs.add &!gradeKV "average" avg
\end{minted}
    \end{onlyenv}%
    \begin{onlyenv}<6>%
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let average db subjects #\tikzcoord{eq1}#=#\tikzcoord{eq2}#
  List.map (KVs.find db) subjects / List.length subjects
let main () =
  let gradeKV = ... in
  let avg = #\tikzcoord{R1}#{| average &gradeKV ["math";"compsci";...] |}#\tikzcoord{R2}# in
  #\tikzcoord{R3}#{| KVs.add &!gradeKV "average" avg |}#\tikzcoord{R4}#
\end{minted}
    \end{onlyenv}%
  \end{onlyenv}%
  \begin{tikzpicture}[remember picture,overlay]%
    \begin{onlyenv}<6>
      \node[code,fit=(eq1) (eq2)] (eq) {};
      \draw[overlay,link,<-] (eq) to[out=60,in=180]
      +(0.5,0.5) node[annot,scale=1.2,anchor=west]
      {No type annotation};
      \node[code,rounded corners=0,color=red,fit=(R1) (R2)] (RA) {};
      \node[code,rounded corners=0,color=yellow,fit=(R3) (R4)] (RB) {};
      \node[annot,scale=1.2] at ($(R4)+(1,-1)$) (AR) {Disjoint regions};
      \draw[link] (AR) to[out=90,in=-40] (RA.-8);
      \draw[link] (AR) to[out=180,in=-40] (RB.-8);
    \end{onlyenv}
  \end{tikzpicture}%
\end{frame}

\section{Inference and constraints}

\begin{frame}{Table of contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections,currentsection]
\end{frame}

% \begin{frame}{Inference and constraints}

%   \begin{enumerate}
%   \item Elaborate regions
%   \item Generate constraints for type, linearity and escape checks
%   \item Solve the constraints
%   \item Simplify the resulting types
%   \end{enumerate}
% \end{frame}

% \begin{frame}[fragile]
%   \frametitle{Inference in action}
%   \begin{onlyenv}<1>%
% \begin{minted}[texcomments=true]{OCaml}
% let average db subjects = 
%   List.map (KVs.find db) subjects / List.length subjects

% let main () =
%   let gradeKV = ... in
%   let avg = average &gradeKV ["math";"compsci";...] in
%   KVs.add &!gradeKV "average" avg
% \end{minted}
%   \end{onlyenv}%
%   \begin{onlyenv}<2>%
% \begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
% let average db subjects =
%   List.map (KVs.find db) subjects / List.length subjects

% let main () =
%   let gradeKV = ... in
%   let avg = #\tikzcoord{R1}#{| average &gradeKV ["math";"compsci";...] |}#\tikzcoord{R2}# in
%   #\tikzcoord{R3}#{| KVs.add &!gradeKV "average" avg |}#\tikzcoord{R4}#
% \end{minted}
%   \end{onlyenv}%
%   \begin{tikzpicture}[remember picture,overlay]%
%     \begin{onlyenv}<2>
%       \node[code,rounded corners=0,color=red,fit=(R1) (R2)] (RA) {};
%       \node[code,rounded corners=0,color=yellow,fit=(R3) (R4)] (RB) {};
%     \end{onlyenv}
%   \end{tikzpicture}%
% \end{frame}

\begin{frame}[fragile,t,shrink]
  \frametitle{Inference in action}
  \begin{center}
    \begin{tikzpicture}[xscale=1,yscale=0.8]
      \node[annot,text width=11.5cm,scale=0.5] at (0,4) (S1) {%
\begin{minted}[texcomments=true]{OCaml}
let average db subjects = 
  List.map (KVs.find db) subjects / List.length subjects

let main () =
  let gradeKV = ... in
  let avg = average &gradeKV ["math";"compsci";...] in
  KVs.add &!gradeKV "average" avg
\end{minted}
      };

      \node[annot,text width=12cm,scale=0.5] at (8,4) (S2) {%
\begin{minted}[texcomments=true,escapeinside=\#\#]{OCaml}
let average db subjects =
  List.map (KVs.find db) subjects / List.length subjects

let main () =
  let gradeKV = ... in
  let avg = #\tikzcoord{R1}#{| average &gradeKV ["math";"compsci";...] |}#\tikzcoord{R2}# in
  #\tikzcoord{R3}#{| KVs.add &!gradeKV "average" avg |}#\tikzcoord{R4}#
\end{minted}
        \tikz[remember picture,overlay]{
          \node[code,rounded corners=0,color=red,fit=(R1) (R2)] (RA) {};
          \node[code,rounded corners=0,color=yellow,fit=(R3) (R4)] (RB) {};
        }%
      };

      \draw[link,->] (S1) to node[midway,above,sloped] {1. Elaborate} (S2) ;
      \draw[link,->] (S1) to node[midway,below,sloped] {regions} (S2) ;

      \visible<2->{
        \node[annot,text width=4.8cm] at (8,0) (S3) {%
          \vspace{-1em}
          \begin{align*}
            \E = &(\tvar_f : \kvar_f)
                 (\tvar_x : \kvar_x)\dots\\
            C = &(\tvar_f \leq \gamma \tarr{\kvar_1} \beta )
                \Cand
                (\gamma \leq \tvar_x)\\
                &\Cand
                (\beta \times  \tvar_x \leq \alpha_r)
                \Cand
                  (\kvar_x \leq \kun)\\
                 &\Cand \dots
          \end{align*}
        };
        \draw[link,->] (S2) to
        node[midway,right,text width=3cm] {2. Generate\\ constraints}
        (S3);
      }

      \begin{onlyenv}<3->
        \node[annot,text width=10.5cm,scale=0.7] at (0.7,1) (S4) {
          @val average : 'k_2 . &(kv_store,'k) -{'k_2}> string list -{'k}> int list@
        };
        \draw[link,->] ($(S3.west)+(0,1)$) to
        node[midway, above, sloped] {3. Solve}
        node[midway, below, sloped] {constraints}
        (S4.east);
      \end{onlyenv}

      \begin{onlyenv}<4->
        \node[annot,text width=9.7cm,scale=0.7] at (0.7,-1) (S5) {
          @val average : &(kv_store,'k) -> string list -{'k}> int list@
        };
        \draw[link,->] (S4) to
        node[midway, left] {4. Simplify}
        node[midway, right] {types}
        (S5);
      \end{onlyenv}

    \end{tikzpicture}
  \end{center}

  \only<1>{%
    1. Infer the placement of region based on the position of borrows and the borrowing rules.
  }%
  \only<2>{%
    2. Generate custom constraints based on HM(X).
  }%
  \only<3>{%
    3. Solve the constraints using a custom algorithm and obtain principal type schemes.
  }%
  \only<4>{%
    4. Simplify the obtained type scheme by leveraging subkinding
  }%

\end{frame}
\section*{Conclusion}

\begin{frame}
  \frametitle{The \lang language -- Summary}

  Prototype: \url{https://affe.netlify.com/}

  \begin{itemize}
  \item[\Y] Linearity, Closures, Borrows and Regions\\
    $\Rightarrow$ Good support for both imperative and functional programming
  \item[\Y] Support managed and unmanaged objects
  \item[\Y] Principal type inference
  \item[\N] No flow sensitivity
  \item[\N] No concurrency story (yet)
  \end{itemize}

  Many examples in the paper: files, session types, semi-persistent arrays, iterators, connection pools, \dots
\end{frame}

\begin{frame}
  \frametitle{The theory -- Summary}

  \begin{itemize}
  \item A Syntax-directed type system for \lang\\
    $\Rightarrow$ How to encode borrows into an ML-style type-system
  \item A formal semantics that takes allocations into account (+ proof of soundness)
  \item An inference algorithm for \lang:
    \begin{itemize}
    \item An extension of HM(X) with kinds
    \item A novel constraint systems to encode linearity and borrows
    \item A constraint solving algorithm, and its proof of completeness
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}[standout]

  \centering\Huge\texttt{Close(Talk)}

\end{frame}

% \bibliography{biblio}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-command-extra-options: "-shell-escape"
%%% TeX-master: "slides-icfp"
%%% End: